---
title: "R Notebook"
output: html_notebook
---


loading libraries..
```{r}
library(lattice)
library(ISLR)
library(MASS)
library(ggplot2)
library(caret)
library(tidyverse)
```



```{r}
nbi <- read.csv('nbi_timerseries.csv')
head(nbi)
```
```{r}
nbi$Maintainer <- as.factor((nbi$Maintainer))
nbi$State.Code <- as.factor((nbi$State.Code))
nbi$Owner <- as.factor((nbi$Owner))
nbi$Material <- as.factor((nbi$Material))
nbi$Structure.Type <- as.factor((nbi$Structure.Type))
head(nbi)
```


```{r}
index = createDataPartition(y=nbi$Category, p=0.7, list=FALSE)

train.set = nbi[index,]
test.set = nbi[-index,]

dim(train.set)
```
```{r}
dim(test.set)
```


```{r}
nbi.tree = train(Category ~ State.Code + Age_L + Material + ADT.Category +  ADTT.Category + Maintainer + Structure.Type,
                 data = train.set,
                 method = "rpart",
                 trControl = trainControl(method = "cv"))
```

```{r}
nbi.tree
```

```{r}
suppressMessages(library(rattle))

fancyRpartPlot(nbi.tree$finalModel)
```

```{r}
nbi.pred = predict(nbi.tree, newdata = test.set)
```
```{r}
table(nbi.pred, test.set$Category)
error.rate = round(mean(nbi.pred != test.set$Category), 2)
error.rate
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

