---
title: "R Notebook"
output: html_notebook
author: Akshay Kale
description: Comparison of decision tree, random forest, and xgb 
---
## Performance Comparison of the Decision Tree, Random Forest, and XGB Model to model deck intervention.
### Import Data files
##### *Reading csv files*
```{r}
dt_mets = read.csv('../../../../../Google\ Drive/Data/trees/metrics/dt/tree-metric-deck-nou.csv') 
rf_mets = read.csv('../../../../../Google\ Drive/Data/trees/metrics/rf/rf-metrics-deck-nou.csv')
rf_imp = read.csv('../../../../../Google\ Drive/Data/trees/metrics/rf/rf-varImp-deck-nou.csv')
```

##### *Preview Decision tree metrics*
```{r}
dt_mets
```

##### *Preview Random forest metrics*
```{r}
rf_mets
```

### Visual comparison of performance meterics between Decision Tree and Random Forest

#### *Kappa vs AUC*
```{r}
library(ggplot2)
# Decision Tree (Kappa vs AUC)
ggplot(dt_mets, aes(x=Kappa, y=AUC)) + 
  geom_point(color='red') + 
  geom_text(label = dt_mets$Tunelength) +
  labs(title="Decision Tree (Kappa vs AUC)")

# Random Forest (Kappa vs AUC)
ggplot(rf_mets, aes(x=Kappa, y=AUC)) + 
  geom_point(color='blue') + 
  geom_text(label = rf_mets$Tunelength ) +
  labs(title="Random Forest (Kappa vs AUC)")
```

#### *Kappa value with respect to tune length
```{r}
# Decision kappa value with respect to tune length
p_kappa_dt <- ggplot(dt_mets, aes(x=Tunelength, y=Kappa)) +
  geom_line() +
  labs(title = "Decision Tree Kappa values with respect to Tunelength")
p_kappa_dt

# Random forest kappa value with respect to tune length
p_kappa_rf <- ggplot(rf_mets, aes(x=Tunelength, y=Kappa)) +
  geom_line() +
  labs(title = "Random forest Kappa values with respect to Tunelength")
```

#### *AUC value with respect to tune length*
```{r}
# Decision Tree kappa value with respect to tune length
p_auc_dt <- ggplot(dt_mets, aes(x=Tunelength, y=AUC)) +
  geom_line() +
  labs(title = "Decision Tree AUC values with respect to Tunelength")
p_auc_dt

# Random forest kappa value with respect to tune length
p_auc_rf <- ggplot(rf_mets, aes(x=Tunelength, y=AUC)) +
  geom_line() +
  labs(title = "Random forest AUC values with respect to Tunelength")
p_auc_rf
```

#### *Preview of the Important Variables*
```{r}
rf_imp_5 <- rf_imp[ which(rf_imp$tunelength==5), ]
rf_imp_10 <- rf_imp[ which(rf_imp$tunelength==10), ]
rf_imp_15 <- rf_imp[ which(rf_imp$tunelength==15), ]
rf_imp_20 <- rf_imp[ which(rf_imp$tunelength==20), ]
rf_imp_25 <- rf_imp[ which(rf_imp$tunelength==25), ]
rf_imp_30 <- rf_imp[ which(rf_imp$tunelength==30), ]
rf_imp_35 <- rf_imp[ which(rf_imp$tunelength==35), ]
rf_imp_40 <- rf_imp[ which(rf_imp$tunelength==40), ]
rf_imp_45 <- rf_imp[ which(rf_imp$tunelength==45), ]
rf_imp_50 <- rf_imp[ which(rf_imp$tunelength==50), ]
rf_imp_55 <- rf_imp[ which(rf_imp$tunelength==55), ]
rf_imp_60 <- rf_imp[ which(rf_imp$tunelength==60), ]
rf_imp_75 <- rf_imp[ which(rf_imp$tunelength==75), ]
rf_imp_80 <- rf_imp[ which(rf_imp$tunelength==80), ]
rf_imp_85 <- rf_imp[ which(rf_imp$tunelength==85), ]
rf_imp_90 <- rf_imp[ which(rf_imp$tunelength==90), ]
rf_imp_95 <- rf_imp[ which(rf_imp$tunelength==95), ]
rf_imp_10 <- rf_imp[ which(rf_imp$tunelength==100), ]
```

```{r}
rf_imp
```
##### *Top 10 important elements*
```{r}
# We can create a dumbbell chart: 
  # 1. Every attribute will have a range of ranks
  # 2. Average of the importance of the variables

library(dplyr)
p <- ggplot(head(rf_imp_5), aes(names, score)) +
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle=45))
p
```
```{r}
library(dplyr)
p <- ggplot(head(rf_imp_10), aes(names, score)) +
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle=45))
p
```

##### *Correlation of the Scores and Variable importance*
```{r}
library(ggpubr)
# Import dataset
df_rf <- read.csv("../../../data/trees/decision-tree-dataset/decision_tree.csv")

# Attributes selection
df_rf_sample <- df_rf %>% select(current.deck, current.superstructure, current.superstructure, total.deck.intervention, total.sub.intervention, total.super.intervention, snowfall, freezethaw, precipitation, score)
```

```{r}
library(corrplot)
library(RColorBrewer)
df_rf_sample <- na.omit(df_rf_sample)
corr <-cor(df_rf_sample)
corrplot(corr, type="upper", order="hclust",col=brewer.pal(n=8, name="RdYlBu"))
```
```{r}
data %>%
  arrange(val) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(name=factor(name, levels=name)) %>%   # This trick update the factor levels
  ggplot( aes(x=name, y=val)) +
    geom_segment( aes(xend=name, yend=0)) +
    geom_point( size=4, color="orange") +
    coord_flip() +
    theme_bw() +
    xlab("")
```

