---
title: "R Notebook"
output: html_notebook
author: Akshay Kale
description: Comparison of decision tree, random forest, and xgb 
---
## Performance Comparison of the Decision Tree, Random Forest, and XGB Model to model deck intervention.
### Import Data files
##### *Reading csv files*
```{r}
dt_mets = read.csv('../../../data/trees/metrics/dt/tree-metric-deck-nou.csv') 
rf_mets = read.csv('../../../data/trees/metrics/rf/rf-metrics-nou.csv')
```

##### *Preview Decision tree metrics*
```{r}
dt_mets
```

##### *Preview Random forest metrics*
```{r}
rf_mets
```

### Visual comparison of performance meterics between Decision Tree and Random Forest

#### *Kappa vs AUC*
```{r}
library(ggplot2)

# Decision Tree (Kappa vs AUC)
ggplot(dt_mets, aes(x=Kappa, y=AUC)) + 
  geom_point(color='red') + 
  geom_text(label = dt_mets$Tunelength) +
  labs(title="Decision Tree (Kappa vs AUC)")

# Random Forest (Kappa vs AUC)
ggplot(rf_mets, aes(x=Kappa, y=AUC)) + 
  geom_point(color='blue') + 
  geom_text(label = rf_mets$Tunelength ) +
  labs(title="Random Forest (Kappa vs AUC)")
```

#### *Kappa value with respect to tune length
```{r}
# Decision kappa value with respect to tune length
p_kappa_dt <- ggplot(dt_mets, aes(x=Tunelength, y=Kappa)) +
  geom_line() +
  labs(title = "Decision Tree Kappa values with respect to Tunelength")
p_kappa_dt

# Random forest kappa value with respect to tune length
p_kappa_rf <- ggplot(rf_mets, aes(x=Tunelength, y=Kappa)) +
  geom_line() +
  labs(title = "Random forest Kappa values with respect to Tunelength")
```

#### *AUC value with respect to tune length*
```{r}
# Decision Tree kappa value with respect to tune length
p_auc_dt <- ggplot(dt_mets, aes(x=Tunelength, y=AUC)) +
  geom_line() +
  labs(title = "Decision Tree AUC values with respect to Tunelength")
p_auc_dt

# Random forest kappa value with respect to tune length
p_auc_rf <- ggplot(rf_mets, aes(x=Tunelength, y=AUC)) +
  geom_line() +
  labs(title = "Random forest AUC values with respect to Tunelength")
p_auc_rf
```